---
title: 'Comparison of 2 approaches for parameter uncertainty'
output: pdf_document
classoption: landscape
author: Sasha D. Hafner
date: "`r format(Sys.time(), '%d %B, %Y %h:%m')`"
---

# Packages

```{r}
library(data.table)
library(knitr)
library(ALFAM2)
library(ggplot2)
```

```{r}
packageVersion('ALFAM2')
```

# Get pars

```{r}
pdl <- fread('../pars/output/pars_boot_long.csv')
pdc <- fread('../pars/output/pars_boot.csv')
ps3 <- fread('../pars/output/pars.csv')
```

```{r}
ps3 <- ps3[pars == 'ps3', ]
nn <- names(ps3)
ps3 <- as.numeric(ps3)
names(ps3) <- nn
ps3 <- ps3[!is.na(ps3)]
```

```{r}
ggplot(pdl, aes(value)) +
  geom_histogram() +
  facet_wrap(~ parameter, scale = 'free')
```

# Sampling function 

```{r}
getPars <- function(pp, n = 1) {

  if (n > 1) {
    pars <- NULL
    for (i in 1:n) {
      p <- data.table::data.table(t(getPars(pp, 1)))
      pars <- rbind(pars, p)
    }
    return(pars)
  }

  pp <- pp[sample(1:nrow(pp), nrow(pp)), ]
  p <- pp[!duplicated(pp[, parameter]), ]
  pars <- as.numeric(p[, value])
  names(pars) <- as.character(p[, parameter])

  return(pars)
}

```

# Get uncorrelated parameters

```{r}
pduc <- getPars(pdl, 100)
```

# Compare

```{r}
round(cor(pdc), 2)
cc <- cor(pdc)
cc <- c(cc)
cc <- cc[cc != 1]
hist(cc)
```

```{r}
round(cor(pduc), 2)
```

# Input data

Vary inputs.

```{r}
n <- 100
set.seed(123)
dat <- data.table(id = 1:n, ct = rep(168, n), app.mthd = sample(c('bc', 'bsth', 'ts', 'os'), n, replace = TRUE), 
                  man.dm = rnorm(n, 5, 1), man.ph = rnorm(n, 7, 0.5), wind.sqrt = rnorm(n, sqrt(5), sqrt(0.5)), 
                  air.temp = rnorm(n, 10, 2), app.rate.ni = rnorm(n, 40, 10), TAN.app = 100)
```

Sort by application method for plotting.

```{r}
dat <- dat[order(app.mthd), ]
dat[, id := 1:nrow(dat)]
```

```{r}
pred <- alfam2(dat, pars = ps3, group = 'id')
dat1 <- merge(dat, pred)
```

```{r}
ggplot(dat1, aes(id, er, colour = app.mthd)) +
  geom_point()
```

# Confidence intervals

```{r}
args(alfam2)
```

With correlated parameters.

```{r}
predcor <- alfam2(dat, group = 'id', conf.int = 0.95, pars.ci = pdc)
datcor <- merge(dat, predcor)
```

Uncorrelated.

```{r}
preduc <- alfam2(dat, group = 'id', conf.int = 0.95, pars.ci = pduc)
datuc <- merge(dat, preduc)
```

```{r}
ggplot(dat1, aes(id, er, colour = app.mthd)) +
  geom_point() +
  geom_errorbar(data = datcor, aes(ymin = er.lwr, ymax = er.upr), alpha = 0.3, width = 0, lwd = 2) +
  geom_errorbar(data = datuc, aes(ymin = er.lwr, ymax = er.upr), alpha = 0.3, width = 2)
```

Error bars with whiskers are from uncorrelated parameters, and wide transparent bars are from the correlated values (complete sets).

# Conclusion
Confidence intervals are always wider when correlation among parameters is ignored.








