---
title: 'Exploration of rate constant/time substitution'
output: pdf_document
classoption: landscape
author: Sasha D. Hafner
date: "`r format(Sys.time(), '%d %B, %Y %h:%m')`"
---

# Packages

```{r}
library(data.table)
library(knitr)
library(ALFAM2)
library(ggplot2)
```

```{r}
packageVersion('ALFAM2')
```

# Setup

Parameters.
Mitigation (0), reference (1), and doubled (2) r1 and r3.

```{r}
print(ALFAM2::alfam2pars03)
p0 <- c(int.f0 = 0.4, int.r1 = -1.5 - 0.3, int.r2 = -1.2, int.r3 = -2.7 - 0.3, int.r5 = -1.8)
p1 <- c(int.f0 = 0.4, int.r1 = -1.5, int.r2 = -1.2, int.r3 = -2.7, int.r5 = -1.8)
p2 <- c(int.f0 = 0.4, int.r1 = -1.5 + 0.3, int.r2 = -1.2, int.r3 = -2.7 + 0.3, int.r5 = -1.8)
```

Input data.

```{r}
dat <- data.table(ct = c(2, 4, 8) * 24, TAN.app = 100)
```

# Predictions

```{r}
pred0 <- alfam2(dat, pars = p0)
pred1 <- alfam2(dat, pars = p1)
pred2 <- alfam2(dat, pars = p2)
```

Doubling pars effect:

```{r}
(pred2 / pred1)[, 'er'] - 1
```

Halving:


```{r}
1 - (pred0 / pred1)[, 'er']
```

Doubling time:

```{r}
pred1[3, 'er'] / pred1[2, 'er'] - 1
```

Halving time:

```{r}
1 - pred1[1, 'er'] / pred1[2, 'er']
```

Apparent mitigation effect at reference time:

```{r}
1 - pred0[2, 'er'] / pred1[2, 'er']
```

At later time.

```{r}
1 - pred0[3, 'er'] / pred1[3, 'er']
```

And under higher emission conditions.

```{r}
1 - pred1[2, 'er'] / pred2[2, 'er']
```

Later:

```{r}
1 - pred1[3, 'er'] / pred2[3, 'er']
```

# Single-pool model

```{r}
p0 <- c(int.f0 = 100, int.r1 = -1.5 - 0.3, int.r2 = -100, int.r3 = -100, int.r5 = -100)
p1 <- c(int.f0 = 100, int.r1 = -1.5      , int.r2 = -100, int.r3 = -100, int.r5 = -100)
p2 <- c(int.f0 = 100, int.r1 = -1.5 + 0.3, int.r2 = -100, int.r3 = -100, int.r5 = -100)
```

```{r}
pred0 <- alfam2(dat, pars = p0)
pred1 <- alfam2(dat, pars = p1)
pred2 <- alfam2(dat, pars = p2)
```

Doubling pars effect:

```{r}
(pred2 / pred1)[, 'er'] - 1
```

Halving:


```{r}
1 - (pred0 / pred1)[, 'er']
```

Doubling time:

```{r}
pred1[3, 'er'] / pred1[2, 'er'] - 1
```

Halving time:

```{r}
1 - pred1[1, 'er'] / pred1[2, 'er']
```

Apparent mitigation effect at reference time:

```{r}
1 - pred0[2, 'er'] / pred1[2, 'er']
```

At later time.

```{r}
1 - pred0[3, 'er'] / pred1[3, 'er']
```

And under higher emission conditions.

```{r}
1 - pred1[2, 'er'] / pred2[2, 'er']
```

Later:

```{r}
1 - pred1[3, 'er'] / pred2[3, 'er']
```

# Conclusions

* Predicted emission is much more sensitive to a fixed relative change in emission rate constants than to time
* But mitigation effects drop in response to increases from either time or emission rate constants, although much more for changes in emission rate constants
* For a single-pool first-order model effects of time and r1 are interchageable
