---
title: 'Test of ALFAM2 closed-form solution'
author: 'Sasha D. Hafner'
date: "`r format(Sys.time(), '%d %B, %Y')`"
urlcolor: blue
---

# Overview
The `alfam2()` function uses a closed-form solution to calculate emission over time.
In this document that solution is checked by comparison to a simple numerical solution.

# Model structure

With a sink for slow pool *S*, structure is given in Fig. 1.

![Structure of ALFAM2 model.](fig1e.pdf)

So derivatives are:

```
df/dt <-  -r1 * F - r2 * F
ds/dt <- r2 * F - r3 * S - r5 * S
de/dt <- r1 * F + r3 * S
```

# Prep

Packages.

```{r}
library(ALFAM2)
```

```{r}
packageVersion('ALFAM2')
```

```{r}
library(data.table)
library(ggplot2)
library(deSolve)
```

# Predictions

Input data--no predictor variables so only intercept parameters are used.

```{r}
dat <- data.table(time = 0:168, TAN_app = 100)
```

Closed-form solution with `alfam2()` function from ALFAM2 package.

```{r}
predscf <- data.table(alfam2(dat, pars = ALFAM2::alfam2pars03_alpha, app.name = 'TAN_app', 
                             time.name = 'time'))
```

Numerical solution uses `lsoda()`.

```{r}
pars <- ALFAM2::alfam2pars03_alpha
pars <- pars[grepl('^int', names(pars))]
pars[grepl('\\.f', names(pars))] <- logistic(pars[grepl('\\.f', names(pars))])
pars[grepl('\\.r', names(pars))] <- 10^(pars[grepl('\\.r', names(pars))])
names(pars) <- gsub('int\\.', '', names(pars))
pars
y <- c(f = pars['f0'] * 100, s = (1 - pars['f0']) * 100, e = 0)
pars <- pars[-1]

rates <- function(t, x, parms) {
  r1 <- parms['r1']
  r2 <- parms['r2']
  r3 <- parms['r3']
  r5 <- parms['r5']

  f <- x[1]
  s <- x[2]

  dfdt <-  -r1 * f - r2 * f
  dsdt <- r2 * f - r3 * s - r5 * s
  dedt <- r1 * f + r3 * s
 return(list(c(dfdt, dsdt, dedt)))

}

predsnum <- lsoda(y = y, times = dat[, time], func = rates, parms = pars)
```

Combine results.

```{r}
dat <- merge(dat, predscf, by = 'time')
dat <- merge(dat, predsnum, by = 'time', suffixes = c('.cf', '.num'))
```

Compare.

```{r}
ggplot(dat, aes(time, e.num)) +
  geom_point() +
  geom_line(aes(y = e.cf))
```

Check difference--very small.

```{r}
dat[, e.diff := e.num - e.cf]
plot(log10(abs(e.diff)) ~ time, data = dat)
```

# Conclusion
The closed-form solution seems to be accurate.
