---
title: 'Test of ALFAM2 closed-form solution'
author: 'Sasha D. Hafner'
date: "`r format(Sys.time(), '%d %B, %Y')`"
urlcolor: blue
---

# Overview
The `alfam2()` function uses a closed-form solution to calculate emission over time.
This document has tests of two closed-form solutions directly without use of the ALFAM2 package.
The solutions are defined here as R functions.
They are tested by comparison to numerical results.

# Model structure

With a sink for slow pool *S*, structure is given in Fig. 1.

![Structure of ALFAM2 model.](fig1e.pdf)

So derivatives are:

```
df/dt <-  -r1 * F - r2 * F
ds/dt <- r2 * F - r3 * S - r5 * S
de/dt <- r1 * F + r3 * S
```

# Prep

Packages.

```{r}
library(data.table)
library(ggplot2)
library(deSolve)
```

# Solution

Solution A, used in package, originally from Valdemar.

```{r}
ALFAMa <- function(t, app, f0, r1, r2, r3, r5) {

  rf <- r1 + r2
  rs <- r3 + r5
  rd <- rf - rs

  fi <- f0 * app
  si <- (1 - f0) * app

  f <- fi * exp(-rf * t)
  s <- exp(-rs * t) * (si - (exp(-rd * t) - 1) / rd * r2 * fi)

  ef <- (1 - exp(-rf * t)) * r1 / rf * fi
  es <- (fi + si - f - s - ef) / rs * r3
  e <- ef + es

  return(data.table(time = t, app = app, f0 = f0, r1 = r1, r2 = r2, 
                    r3 = r3, r5 = r5, f = f, s = s, e = e))
}
```

Solution B, from Paul.

```{r}
ALFAMb <- function(t, app, f0, r1, r2, r3, r5) {

  rf <- r1 + r2
  rs <- r3 + r5
  rd <- rf - rs
  fi <- f0 * app
  si <- (1 - f0) * app

  f <- fi * exp(-rf * t)
  s <- (si - (exp(-rd * t) - 1) * r2 * fi / rd) * exp(-rs * t)

  ef <- fi * (r1 / rf) * (1 - exp(-rf * t))
  es <- (si + fi * r2 / rd) * (r3 / rs) * (1 - exp(-rs * t)) - 
        (fi * r2 / rd) * (r3 / rf) * (1 - exp(-rf * t))

  e <- ef + es

  return(data.table(time = t, app = app, f0 = f0, r1 = r1, r2 = r2, 
                    r3 = r3, r5 = r5, f = f, s = s, e = e))
}
```


# Predictions

Input data.

```{r}
dat <- data.table(time = 0:168)
```

Parameter values.

```{r}
r1 <- 0.03
r2 <- 0.01
r3 <- 0.001
r5 <- 0.05
f0 <- 0.3
```


Closed-form solution A.

```{r}
predcfa <- ALFAMa(t = dat[, time], app = 100, f0, r1, r2, r3, r5)
```

And B.

```{r}
predcfb <- ALFAMb(t = dat[, time], app = 100, f0, r1, r2, r3, r5)
```

Numerical solution uses `lsoda()`.

```{r}
pars <- c(r1 = r1, r2 = r2, r3 = r3, r5 = r5, f0 = f0)
app <- 100
y <- c(f = pars['f0'] * app, s = (1 - pars['f0']) * app, e = 0)
names(y) <- c('f', 's', 'e')

rates <- function(t, x, parms) {
  r1 <- parms['r1']
  r2 <- parms['r2']
  r3 <- parms['r3']
  r5 <- parms['r5']

  f <- x[1]
  s <- x[2]

  dfdt <- as.numeric(-r1 * f - r2 * f)
  dsdt <- as.numeric(r2 * f - r3 * s - r5 * s)
  dedt <- as.numeric(r1 * f + r3 * s)

  return(list(c(dfdt, dsdt, dedt)))

}

prednum <- data.table(lsoda(y = y, times = dat[, time], func = rates, parms = pars))
```

Combine results.

```{r}
pred <- merge(predcfa, prednum, by = 'time', suffixes = c('', '.num'))
pred <- merge(pred, predcfb, by = 'time', suffixes = c('.cfa', '.cfb'))
```

Differences.

```{r}
pred[, diffa := e.cfa - e.num]
pred[, diffb := e.cfb - e.num]
```

Compare.

```{r}
plot(e.num ~ time, data = pred)
points(e.cfa ~ time, data = pred, col = 'red', cex = 0.7)
points(e.cfb ~ time, data = pred, col = 'blue', cex = 0.4)
```

```{r}
plot(f.num ~ time, data = pred)
points(f.cfa ~ time, data = pred, col = 'red', cex = 0.7)
points(f.cfb ~ time, data = pred, col = 'blue', cex = 0.4)
```

```{r}
plot(s.num ~ time, data = pred)
points(s.cfa ~ time, data = pred, col = 'red', cex = 0.7)
points(s.cfb ~ time, data = pred, col = 'blue', cex = 0.4)
```

Difference from numerical solution.

```{r}
plot(log10(abs(diffa)) ~ time, data = pred, col = 'red')
points(log10(abs(diffb)) ~ time, data = pred, col = 'blue')
```

# Conclusion
Both closed-form solutions seem to be accurate.

# Something else: try to blow up solution

```{r}
 ALFAMa(t = 168, app = 100, f0 = 0.5, r1 = 0.1, r2 = 0.1, r3 = 0.01, r5 = 0.01)
 ALFAMa(t = 168, app = 100, f0 = 0.5, r1 = 1E7, r2 = 0.1, r3 = 0.01, r5 = 0.01)
 ALFAMa(t = 168, app = 100, f0 = 0.5, r1 = 0.1, r2 = 1E7, r3 = 0.01, r5 = 0.01)
 ALFAMa(t = 168, app = 100, f0 = 0.5, r1 = 0.1, r2 = 0.1, r3 = 1E7, r5 = 0.01)
 ALFAMa(t = 168, app = 100, f0 = 0.5, r1 = 0.1, r2 = 0.1, r3 = 0.01, r5 = 1E7)
```

Problems are with r3 and r5, presumably through exp(... + rd).

Paul's solution will still give emission because it does not use `s` for calculation of `e`:

```{r}
 ALFAMb(t = 168, app = 100, f0 = 0.5, r1 = 0.1, r2 = 0.1, r3 = 0.01, r5 = 0.01)
 ALFAMb(t = 168, app = 100, f0 = 0.5, r1 = 1E7, r2 = 0.1, r3 = 0.01, r5 = 0.01)
 ALFAMb(t = 168, app = 100, f0 = 0.5, r1 = 0.1, r2 = 1E7, r3 = 0.01, r5 = 0.01)
 ALFAMb(t = 168, app = 100, f0 = 0.5, r1 = 0.1, r2 = 0.1, r3 = 1E7, r5 = 0.01)
 ALFAMb(t = 168, app = 100, f0 = 0.5, r1 = 0.1, r2 = 0.1, r3 = 0.01, r5 = 1E7)
```

A work-around for Valdemar's solution:

```{r}
ALFAMa2 <- function(t, app, f0, r1, r2, r3, r5) {

  rf <- r1 + r2
  rs <- r3 + r5
  rd <- rf - rs

  fi <- f0 * app
  si <- (1 - f0) * app

  erdt <- exp(-rd * t)
  if (erdt > .Machine$double.xmax) erdt <- .Machine$double.xmax / 10

  f <- fi * exp(-rf * t)
  s <- exp(-rs * t) * (si - (erdt - 1) / rd * r2 * fi)

  ef <- (1 - exp(-rf * t)) * r1 / rf * fi
  es <- (fi + si - f - s - ef) / rs * r3
  e <- ef + es

  return(data.table(time = t, app = app, f0 = f0, r1 = r1, r2 = r2, 
                    r3 = r3, r5 = r5, f = f, s = s, e = e))
}
```

```{r}
 ALFAMa2(t = 168, app = 100, f0 = 0.5, r1 = 0.1, r2 = 0.1, r3 = 0.01, r5 = 0.01)
 ALFAMa2(t = 168, app = 100, f0 = 0.5, r1 = 1E7, r2 = 0.1, r3 = 0.01, r5 = 0.01)
 ALFAMa2(t = 168, app = 100, f0 = 0.5, r1 = 0.1, r2 = 1E7, r3 = 0.01, r5 = 0.01)
 ALFAMa2(t = 168, app = 100, f0 = 0.5, r1 = 0.1, r2 = 0.1, r3 = 1E7, r5 = 0.01)
 ALFAMa2(t = 168, app = 100, f0 = 0.5, r1 = 0.1, r2 = 0.1, r3 = 0.01, r5 = 1E7)
```

